Identification of Direct Cherenkov Pixels using Boosted Decision Trees

### 2016-05-18 report/bdtarticle.tex
Backup 18/05/16

At present, the DC pixel candidate can be identified by applying a number of cuts to pixels in an image, such as those used by the HESS collaboration \cite{hess07}. The variable $Q_{DC}$ is defined as the ratio of the largest neighbouring pixel intensity to the intensity of a given pixel, and from the subset of pixels passing these cuts, the DC candidate is simply the pixel with the smallest $Q_{DC}$. However, this method is neither reliable nor efficient. An better method would aim to increase the number of correctly identified events, and enable cuts which better discriminate between correctly and incorrectly identified events.

@@

The candidates were checked against the true DC pixels identified in the EAS-free images. Of 2000 testing events it was found that there were 6000 triggered images. Of these images, 5.7\% passed the required cuts. The $Q_{DC}$ was found to be 68\% accurate in identifying the DC pixel in those passing images, as shown in \ref{fig:cutdistribution}. These values served as a benchmark for BDT performance.

The BDT was trained with the Scikit Learn Python package \cite{scikit-learn}. A set of 2000 CORSIKA events was selected, and randomly split through use of the Python random.random() function into the two subsets for training and testing. Within the subset of training events, every triggered HESS 1 image was used. For each of the 3.6 million triggered image pixels, an entry was formed of the variables listed in table \ref{tab:table2}. A score of 0 was assigned to every pixel to indicate background, with the exception of those that were identified as the true DC pixel through the Shower-Free simulation. These pixels were included with a score of 1, identifying it as a signal pixel. Having created a dataset, the BDT was then trained with a maximum depth of 5, and 100 trees generated.

@@

The relative importance of each \textquoteleft feature' is automatically calculated by the Scikit Learn package, and is also recorded in table \ref{tab:table2}. The variable $DC_{Count}$ was consistently the most importance variable across many combinations of included variables and BDT training parameters. It was found that, under the conditions listed above, the BDT was 99.94 \% accurate for the entire training dataset, and 99.93 \%  accurate for the testing dataset. This indicates that the BDT was not significantly overtrained, which would otherwise be manifested by a large divergence in accuracy between testing and training data.

Having trained the BDT successfully, it was then applied to the same separate dataset as for the classic QDC identification. In each camera image, the event with the largest BDT score was deemed to be \textquoteleft most signal-like', and thus selected as the DC pixel candidate. A cut was applied, requiring $P_{signal} > 0.5$ for the DC candidate to be accepted. A second cut requiring $DC_{Count} > 200$ removed many incorrectly identified events.

Application of this combined cut greatly increases the successful identification rate. Of the 2000 events, 57\% of events passed all of the required cuts. The BDT was found to be 85 \% accurate in identifying the DC pixel in those passing events. This represents a very significant improvement in identification accuracy over the previous QDC method, and corresponds to a fivefold increase in the number of HESS telescope images with corresponding identified DC pixels.

>>>>

At present, the DC pixel candidate can be identified by applying a number of cuts to pixels in an image, such as those used by the HESS collaboration \cite{hess07}. The variable $Q_{DC}$ is defined as the ratio of the largest neighbouring pixel intensity to the intensity of a given pixel, and from the subset of pixels passing these cuts, the DC candidate is simply the pixel with the smallest $Q_{DC}$. Despite a good degree of accuracy, this method is very inefficient, leaving the majority of events having no DC candidate from the $Q_{DC}$ method. A better method would aim to increase the number of correctly identified events, while still enabling cuts which discriminate well between correctly and incorrectly identified events.

@@

The candidates were checked against the true DC pixels identified in the EAS-free images. Of 2000 testing events it was found that there were 6000 triggered images. Of these images, 5.2\% passed the required cuts. The $Q_{DC}$ was found to be 86.2\% accurate in identifying the DC pixel in those passing images, as shown in \ref{fig:cutdistribution}. These values served as a benchmark for BDT performance.

The BDT was trained with the Scikit Learn Python package \cite{scikit-learn}. Using the training set of 2000 CORSIKA events, and randomly split through use of the Python random.random() function into two further subsets, one for learning and one to check overtraining. Within the subset of learning events, every HESS 1 image was used, provided it was triggered in both EAS-free and full-shower simulations. For each of the 1.7 million triggered image pixels, an entry was formed of the variables listed in table \ref{tab:table2}. A class of 0 was assigned to every non-DC pixel, and a class of 1 was assigned to every DC pixel. Having created a dataset, the BDT was then trained with a maximum depth of 8, and 100 trees generated.

@@

The relative importance of each \textquoteleft feature' is automatically calculated by the Scikit Learn package, and is also recorded in table \ref{tab:table2}. The variable $DC_{Count}$ was consistently the most importance variable across many combinations of included variables and BDT training parameters. It was found that, under the conditions listed above, the BDT was 99.94 \% accurate for the entire learning pixel set, and 99.93 \%  accurate for the overtraining-check pixel set. This indicates that the BDT was not significantly overtrained, which would otherwise be manifested by a large divergence in accuracy between learning and overtraining-check data.

Having trained the BDT successfully, it was then applied to the same test dataset as for the classic QDC identification. In each camera image, the event with the largest BDT score was deemed to be \textquoteleft most signal-like', and thus selected as the DC pixel candidate. A cut was applied, requiring $P_{signal} > 0.5$ for the DC candidate to be accepted. A second cut requiring $DC_{Count} > 150$ removed many incorrectly identified events.

Application of this combined cut greatly increases the successful identification rate. Of the 2000 events, 37.8\% of images passed all of the required cuts. The BDT was found to be 86.0\% accurate in identifying the DC pixel in those passing events. This represents a very significant improvement in pixel identification efficiency, at the cost of a negligible increase in the fraction of incorrectly identified pixels.
